---
title: "visualize"
author: "Marina Botelho"
date: "12/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read data

```{r}
# libraries
library(here)
library(readr)
library(DT)

# variables
url_ac <- "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_AC.csv"
# if ERDDAP server down (Error in download.file) with URL above, use this:
#    url_ac <- "https://raw.githubusercontent.com/noaa-iea/r3-train/master/data/cciea_AC.csv"
csv_ac <- here("data/cciea_AC.csv")

# download data
if (!file.exists(csv_ac))
  download.file(url_ac, csv_ac)

# read data
d_ac        <- read_csv(csv_ac, col_names = F, skip = 2)
names(d_ac) <- names(read_csv(csv_ac))
#difference between read_csv (from readr) and read-csv:
#Copy from Ben Best explanation: "I prefer readr::read_csv() over read.csv() since columns of character type are not converted to type factor by default. It will also default to being read in as a tibble rather than just a data.frame."   

# show data
datatable(d_ac)
```

## Plot statically with ggplot2

### Simple line scan (geom_line())

```{r}
library(dplyr)
library(ggplot2)

# subset data
d_coast <- d_ac %>% 
  # select columns
  select(time, total_fisheries_revenue_coastwide) %>% 
  # filter rows
  filter(!is.na(total_fisheries_revenue_coastwide))
datatable(d_coast)
```

```{r}
# ggplot object
p_coast <- d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = time, y = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_line()

# show plot
p_coast
```

### Trend line (geom_smooth())

```{r}
p_coast + 
  geom_smooth(method = "lm")

#changing method:
p_coast_different <- d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = time, y = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_line() + 
  geom_smooth(method = NULL)
p_coast_different

```

### Distribution of values (geom_histogram())

```{r}
d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_histogram(binwidth=25)
```

### Series line plot (aes(color=region))

```{r}
library(stringr)
library(tidyr)

d_rgn <- d_ac %>% 
  # select columns
  select(
    time, 
    starts_with("total_fisheries_revenue")) %>% 
  # exclude column
  select(-total_fisheries_revenue_coastwide) %>% 
  # pivot longer
  pivot_longer(-time) %>% 
  # mutate region by stripping other
  mutate(
    region = name %>% 
      str_replace("total_fisheries_revenue_", "") %>%
      #upper case:
      str_to_upper()) %>% 
  # filter for not NA
  filter(!is.na(value)) %>% 
  # select columns
  select(time, region, value)
  
# create plot object
p_rgn <- ggplot(
  d_rgn,
  # aesthetics
  aes(
    x     = time, 
    y     = value,
    group = region,
    color = region)) +
  # geometry
  geom_line()

# show plot
p_rgn
```

### Update labels (labs())

```{r}
p_rgn <- p_rgn +
  labs(
    title  = "Fisheries Revenue",
    x      = "Year",
    y      = "Millions $ (year 2015)",
    color  = "Region")
p_rgn
```

### Multiple plots with facet_wrap()

```{r}
p_rgn +
  facet_wrap(vars(region))
```

### Bar plot (geom_col())

```{r}
library(glue)
library(lubridate)

yr_max <- year(max(d_rgn$time))

d_rgn %>% 
  # filter by most recent time
  filter(year(time) == yr_max) %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_col() +
  # add labels
  labs(
    title  = glue("Fisheries Revenue for {yr_max}"),
    x      = "Region",
    y      = "Millions $ (year 2015)",
    fill   = "Region")
```

### Variation of series (geom_boxplot())

```{r}
d_rgn %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_boxplot() +
  # add labels
  labs(
    title  = "Fisheries Revenue Variability",
    x      = "Region",
    y      = "Millions $ (year 2015)") +
  # drop legend since redundant with x axis
  theme(
    legend.position = "none")
```

### Variation of series (geom_violin())

```{r}
p_rgn_violin <- d_rgn %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_violin() +
  # add labels
  labs(
    title  = "Fisheries Revenue Variability",
    x      = "Region",
    y      = "Millions $ (year 2015)") +
  # drop legend since redundant with x axis
  theme(
    legend.position = "none")
p_rgn_violin
```

### Change theme (theme())

```{r}
p_rgn_violin +
  theme_classic()
```

## Plot interactively with ploty or dygraphs

### Make ggplot interactive with plotly::ggplotly()

```{r}
#When rendering to HTML, you can render most ggplot objects interactively with plotly::ggplotly(). The plotly library is an R htmlwidget providing simple R functions to render interactive JavaScript visualizations.

plotly::ggplotly(p_rgn)
```

### Create interactive time series with dygraphs::dygraph()

```{r}
#Another htmlwidget plotting library written more specifically for time series data is dygraphs. Unlike the ggplot2 data input, a series is expected in wide (not tidy long) format. So we use tidyr’s pivot_wider() first.

library(dygraphs)

d_rgn_wide <- d_rgn %>% 
  mutate(
    Year = year(time)) %>% 
  select(Year, region, value) %>% 
  pivot_wider(
    names_from  = region,
    values_from = value)
datatable(d_rgn_wide)
```

```{r}
d_rgn_wide %>% 
  dygraph() %>% 
  dyRangeSelector()
```
